\chapter{Future work}
\label{chap:futurework}

As this thesis covered rather broad area, several aspects will need to be looked in more detail. First of all, it would be interesting to see a similar study being done on a different user group, and also with a longer application testing period with more participants. Similar studies with other user groups are needed in order to be able to generalize the overall design of the application ti fit a wider and more general user group.

It would also be very useful too see a similar study being conducted on a larger task management application such as a calendar, rather than a to-do list. This would require some more work, as a calendar system is naturally larger and more complex than a simple to-do list, but by discovering that context information adds great value to a calendar system would be a big step towards making context valuable for all task management systems.

In our proof of concept application we decided that a recommender would be a good way to test the usefulness of context information in a to-do list. This was used to suggest tasks for the user. Investigating other ways of utilizing contexts in task management systems would also be useful. For example, one could use contexts to keep a calendar very much up to date with the users events. By learning the users behavior, the calendar could create entries in the calendar if the user has not already done so. The calendar could infer the user's activity by looking at contexts. 